Speed estimation for vehicles using \\SVM classification and edge detection

### 2016-06-16 Administratie/report/paper.tex
Abstract tweak
Finale versie paper

\begin{abstract}
The goal of this paper it to determine the maximum speed for a moving vehicle in traffic, using only camera images and information from the steering column. We construct a system that combines SVM classification and edge detection to foresee possible obstacles on the given trajectory of the car. The maximum speed is computed so that braking in time will avoid collision with those obstacles. 
TODO: hoe goed is dit systeem?
\keywords{SVM classification, Local Binary Patterns, edge detection}
\end{abstract}

@@

An additional feature that can be realized using the electronics that are already available in these cars, is automated maximum speed detection. This speed detection would simulate the braking distance towards the nearest obstacle and inform the user. There are various ways to achieve this. Commonly used sensors in adaptive cruise control are lidar and radar.  The choice of sensor presents classic design trade-offs \cite{Jones2001}.  Lidar  is  less  expensive  to  produce  and  easier  to package  but  performs  poorly  in  rain  and  snow. Another problem is that accumulations of mud, dust or snow on the car can block lidar beams. Radar-based  systems,  on  the  other  hand,  can ''see'' at least 150 meters ahead in fog or rain heavy enough to cut the driver's ability to see down to 10 meters or less.

@@

Basically, we want the hyperplane that maximizes the geometric distance to the closest data points (as shown in figure \ref{fig:hyperplane}).

@@

One method for comparing results for a set of parameter values is cross-validation. In $v$-fold cross-validation, we first divide the training set into $v$ subsets of equal size. Sequentially one subset is tested using the classifier trained on the remaining $v − 1$ subsets. Thus, each instance of the whole training set is predicted once so the cross-validation accuracy is the percentage of data which are correctly classified.

@@

\subsection{Training the SVM classifier}
For each combination of the previously mentioned parameters, the SVM classifier must be trained. For a training set (a subset of the datasets available) each frame is divided into blocks. Each block must be manually labeled with the expected value for classification. For practical reasons we put a margin between each block in the frames of the training set and only used the $i$-th frame of each dataset ($i=0,10,...,40$).
Figure \ref{fig:frame-training} shows a visualization of pre-labeled blocks in a frame.
The classifier is trained using the label and calculated feature vector for each block in each frame of the training set.

@@

\subsection{Road marks}
The training dataset was labeled twice, once considering blocks containing road marks as part of the road class and once as part of the non-road class. Initially road marks were labeled as non-road, in order to get a good classification of road (i.e. different types of asphalt). But as these road marks aren't really considered an obstacle for a car driver, ideally they should be classified as road. This makes up for a more difficult classification problem. To quantify this effect we apply classification both with and without road marks.

@@

Different combinations of training sets were selected of the four given datasets (training sets used are each time shown on the x-axis on the graphs). The trained classifier is then tested on the remaining datasets.

All of these results for the classifier can be found in appendix \ref{appendix}.

@@

\subsubsection{Block size}
In general, block sizes 16x16px and 32x32px tend to show the better results. As an example we present the F-scores for using color feature vectors with road marks in figure \ref{fig:col-wi-rm}. Other combinations follow the same trend, although 32x32px sometimes performs slightly better. This is why we conclude the block size of 32x32px to be a more reliable choice.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig/RBF-COL-wi-RM.pdf}
\caption{F-scores for color with road marks using RBF kernel}
\label{fig:col-wi-rm}
\end{figure}

\subsubsection{Feature vector values}
The use of LBP represents no significant gains in order to get a viable classifier for road detection. If we compare results (block size 32x32px) of feature vectors on color only to the combination of LBP and color,  performance remains roughly the same.
Therefore, in the case of our classification problem, we do not add LBP to the feature vector.

@@

\subsubsection{Road marks}
Including road marks as part of the road gets better results for the classification problem.
For our application this is more practical, because road marks should not be considered to be obstacles.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig/with-vs-without-roadmarks.pdf}
\caption{F-scores for 16x16px blocks with color using RBF kernel}
\label{fig:col-wi-rm}
\end{figure}

Considering the results of the different scenarios, we can conclude that training on the fourth dataset results in poorer classification overall. This is because another type of road is used in this dataset, conflicting with the more common road type present in the other datasets. This distorts the classifier increases misclassifications.

\noindent In retrospect, the optimal parameters for our classification problem are:
\begin{itemize}
\item RBF kernel
\item Block size 32x32px
\item Using only color histogram values for the feature vectors
\item Including road marks as part of the road  
\end{itemize}

@@

The SVM classification shows optimal results when using color histogram values for the feature vector. These feature vectors are calculated for each block of 32x32px in an image.
These results tend to get better when we include road marks as part of the road in the training data. This is also optimal for use in our application as we don't want to detect road marks as obstacles.

@@

In order to combine the SVM classification with the edge detection, two strategies were proposed. The first strategy simply intersects the SVM classification results and edges to detect the nearest obstacle in the trajectory. This method show to be the safe option, by often falsely detecting an obstacle.  Results show that the best strategy is to first detect all the edges and then classify these edges as being part of the road or not, using the SVM classifier. Only edges which are not considered part of the road remain. The maximum speed is then calculated based on the nearest edge in the trajectory of the car.

>>>>

\begin{abstract}
The goal of this paper it to determine the maximum speed for a moving vehicle in traffic, using only camera images and information from the steering column. We construct a system that combines SVM classification and edge detection to foresee possible obstacles on the given trajectory of the car. The maximum speed is computed so that braking in time will avoid collision with those obstacles. Several strategies are proposed and compared and we conclude which strategy shows optimal performance for our application.

\keywords{SVM classification, Local Binary Patterns, edge detection}
\end{abstract}

@@

An additional feature that can be realized using the electronics that are already available in these cars, is automated maximum speed detection. This speed detection would simulate the braking distance towards the nearest obstacle and inform the user.

There are various ways to achieve this. Commonly used sensors in adaptive cruise control are lidar and radar.  The choice of sensor presents classic design trade-offs \cite{Jones2001}.  Lidar  is  less  expensive  to  produce  and  easier  to package  but  performs  poorly  in  rain  and  snow. Another problem is that accumulations of mud, dust or snow on the car can block lidar beams. Radar-based  systems,  on  the  other  hand,  can ''see'' at least 150 meters ahead in fog or rain heavy enough to cut the driver's ability to see down to 10 meters or less.

@@

Basically, we want the hyperplane that maximizes the geometric distance to the closest data points (as shown in figure \ref{fig:hyperplane}). With this hyperplane, we can determine for any given data point on which side of the hyperplane it is located, i.e the class the data point belongs to.

@@

One method for comparing results for a set of parameter values is cross-validation. In $v$-fold cross-validation, we first divide the training set into $v$ subsets of equal size. Sequentially one subset is tested using the classifier trained on the remaining $v-1$ subsets. Thus, each instance of the whole training set is predicted once so the cross-validation accuracy is the percentage of data which are correctly classified.

@@

\subsection{Training the SVM classifier}
For each combination of the previously mentioned parameters, the SVM classifier must be trained. For a training set (a subset of the datasets available) each frame is divided into blocks. Each block must be manually labelled with the expected value for classification. For practical reasons we put a margin between each block in the frames of the training set and only used the $i$-th frame of each dataset ($i=0,10,...,40$).
Figure \ref{fig:frame-training} shows a visualization of pre-labelled blocks in a frame.
The classifier is trained using the label and calculated feature vector for each block in each frame of the training set.

@@

\subsection{Road marks}
The training dataset was labelled twice, once considering blocks containing road marks as part of the road class and once as part of the non-road class. Initially road marks were labelled as non-road, in order to get a good classification of road (i.e. different types of asphalt). But as these road marks aren't really considered an obstacle for a car driver, ideally they should be classified as road. This makes up for a more difficult classification problem. To quantify this effect we tested and compared classification results both with and without road marks.

@@

Different combinations of training sets were selected of the four given datasets (training sets used are each time shown on the x-axis on the graphs). The trained classifier is then tested on the remaining datasets. All of these results for the classifier can be found in appendix \ref{appendix}.

@@

\subsubsection{Block size}
In general, block sizes 16x16px and 32x32px tend to show the better results. As an example we present the F-scores for using color feature vectors with road marks in figure \ref{fig:col-wi-rm}. Other combinations follow the same trend. In practice, smaller blocks are more convenient for our application. Smaller blocks allow for a more precise estimation of road, and its edges. This is why we conclude the block size of 16x16px to be a more reliable choice.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig/RBF-COL-wi-RM.pdf}
\caption{F-scores for color with road marks using RBF kernel}
\label{fig:col-wi-rm}
\end{figure}

\subsubsection{Feature vector values}
The use of LBP represents no significant gains in order to get a viable classifier for road detection. If we compare results (block size 16x16px) of feature vectors on color only to the combination of LBP and color,  performance remains roughly the same.
Therefore, in the case of our classification problem, we do not add LBP to the feature vector.
It should be mentioned that we only used the basic form of LBP, without any improvements such as Rotation Invariance \cite{Guo2010}.

@@

\subsubsection{Road marks}
Including road marks as part of the road gets better results for the classification problem, as is shown in figure \ref{fig:wi-vs-wo-RM}.
For our application this is more practical, because road marks should not be considered to be obstacles.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig/with-vs-without-roadmarks.pdf}
\caption{F-scores for 16x16px blocks with color using RBF kernel}
\label{fig:wi-vs-wo-RM}
\end{figure}

Considering the results of the different scenarios, we can conclude that training on the fourth dataset results in poorer classification overall. This is because another type of road is used in this dataset, conflicting with the more common road type present in the other datasets. This distorts the classifier and increases misclassifications.

\noindent In retrospect, the optimal parameters for our classification problem are:
\begin{itemize}
\item RBF kernel
\item Block size 16x16px
\item Using only color histogram values for the feature vectors
\item Including road marks as part of the road  
\end{itemize}

@@

The SVM classification shows optimal results when using color histogram values for the feature vector. These feature vectors are calculated for each block of 16x16px in an image.
These results tend to get better when we include road marks as part of the road in the training data. This is also optimal for use in our application as we don't want to detect road marks as obstacles.

@@

In order to combine the SVM classification with the edge detection, two strategies were proposed. The first strategy simply intersects the SVM classification results and edges to detect the nearest obstacle in the trajectory. This method shows to be the safe option, by often falsely detecting an obstacle.  Results show that the best strategy is to first detect all the edges and then classify these edges as being part of the road or not, using the SVM classifier. Only edges which are not considered part of the road remain. The maximum speed is then calculated based on the nearest edge in the trajectory of the car.
